Analyzing HAC Reduction Penalty Likelihood Using Logistic Regression
========================================================

```{r Data Prep, echo=FALSE}
setwd("~/R/Regression Example")
library(RODBC)

#Connect to SQL Databases and read in data
con <- odbcConnect("AHA",uid="SASUser", pwd="EHI1301a")
aha <- sqlQuery(con,"select t1.ID,MNAME, MCRNUM, MAPP8, MAPP5, t2.HOSPBD  from AS12DEM t1 inner join as12svc1 t2 on t1.id = t2.id")
odbcClose(con)
rm(con)

con <- odbcConnect("MemberList",uid="SASUser", pwd="EHI1301a")
mem <- sqlQuery(con,"select * from AEH_Current_Members")
odbcClose(con)
rm(con)

con <- odbcConnect("IPPSData",uid="SASUser", pwd="EHI1301a")
cmi <- sqlQuery(con,"select PROVIDER_NUMBER, HOSPITAL, TACMIV31 from FY2014IMPACT")
odbcClose(con)
rm(con)

#Read in HAC_Data.csv; rename matching variables in data frames
names(aha) <- c("ID","MNAME","PROVIDER_NUMBER", "MAPP8", "MAPP5", "HOSPBD")
aha$bed.size <- ifelse(aha$HOSPBD >= 400,1,0)

hac <- read.csv("HAC_Data.csv")
names(hac) <- c("PROVIDER_NUMBER","PENALIZED", "SCORE")

mem <- as.data.frame(mem[,c("MCRNUM","MNAME")])
names(mem) <- c("PROVIDER_NUMBER","MNAME")


#Merge data frames
cmi_mem <- merge(cmi, mem, by="PROVIDER_NUMBER")
hac_mem <- merge(hac, mem, by="PROVIDER_NUMBER") 
cmi_hac <- merge(hac, cmi, by="PROVIDER_NUMBER")
cmi_hac_aha <- merge(cmi_hac, aha, by="PROVIDER_NUMBER")
cmi_hac_mem <- merge(cmi_hac, mem, by="PROVIDER_NUMBER")
cmi_hac_aha_mem <- merge(cmi_hac_mem,aha, by="MNAME")

#Recode PENALIZED Variable
cmi_hac_aha$PENALIZED.BINARY <- ifelse(cmi_hac_aha$PENALIZED == "Y",1,0)
cmi_hac_aha_mem$PENALIZED.BINARY <- ifelse(cmi_hac_aha_mem$PENALIZED == "Y",1,0)

#Recode MAPP8 and MAPP5
cmi_hac_aha$MAPP8.BINARY <- ifelse(cmi_hac_aha$MAPP8 == 2,0,1)
cmi_hac_aha_mem$MAPP8.BINARY <- ifelse(cmi_hac_aha_mem$MAPP8 == 2,0,1)
cmi_hac_aha$MAPP5.BINARY <- ifelse(cmi_hac_aha$MAPP5 == 2,0,1)
cmi_hac_aha_mem$MAPP5.BINARY <- ifelse(cmi_hac_aha_mem$MAPP5 == 2,0,1)

```

## Background

As an addition to the hospital readmissions reduction program (HRRP) and the value-based purchasing program (VBP), the third of the Affordable Care Act's quality incentive programs, the hospital acquired-condition (HAC) reduction program, is set to hit hospitals beginning in FY 2015...

Recently released preliminary data from CMS' FY2015 Inpatient Prospective Payment Proposed rule indicate that `r nrow(hac[which(hac$PENALIZED == "Y"),])` hospitals will be penalized under the program with a one percent reduction in hospital payments from the Centers for Medicare & Medicaid Services...

## Analysis

The Essential Hospitals Institute... 

```{r Chi^2 Test}
table(cmi_hac_aha$PENALIZED.BINARY, cmi_hac_aha$MAPP8.BINARY)
chisq.test(cmi_hac_aha$PENALIZED.BINARY, cmi_hac_aha$MAPP8.BINARY)

cmi_hac_aha$MEMBER <- ifelse(cmi_hac_aha$PROVIDER_NUMBER %in% mem$PROVIDER_NUMBER,1,0)
chisq.test(cmi_hac_aha$PENALIZED.BINARY, cmi_hac_aha$MEMBER)
table(cmi_hac_aha$PENALIZED.BINARY, cmi_hac_aha$MEMBER)

```




```{r Regression Model, tidy=TRUE}
logit.model <- glm(PENALIZED.BINARY ~ ADC + MCR_PCT + DSHPCT + MAPP8.BINARY + MAPP5.BINARY, family=binomial(logit), data=cmi_hac_aha)
summary(logit.model)

x <- with(logit.model, null.deviance - deviance)

plot(logit.model$fitted)

round(1-pchisq(173.3525, df = 5),4)

library(popbio)
logi.hist.plot(cmi_hac_aha$DSHPCT,cmi_hac_aha$PENALIZED.BINARY,boxp=FALSE,type="hist",col="gray")

```


